Hexagonal binning algorithm

### 2024-05-14 supplementary/hexagon_binning_algorithm.qmd
Update hexbin algorithm with buffer: q

```{r}
library(ggplot2)
library(dplyr)
library(quollr)
library(readr)

@@

```{r}
#| label: fig-hex-param
#| fig-cap: "Hexagon binning parameters."

@@

@@

Points are allocated to the bin that they fall in by considering the nearest centroid.

```{r}
emb_df <- tibble::tibble(emb1 = 0.774, emb2 = 0.5055) |>
  add_row(emb1 = 0.424, emb2 = 0.5055) |>
  add_row(emb1 = 0.5115, emb2 = 0.3545) |>
  add_row(emb1 = 0.5115, emb2 = 0.6565) |>
  add_row(emb1 = 0.6865, emb2 = 0.3545) |>
  add_row(emb1 = 0.6865, emb2 = 0.6565) |>
  add_row(emb1 = 0.599, emb2 = 0.304) |>
  add_row(emb1 = 0.599, emb2 = 0.707) |>
  add_row(emb1 = 0.424, emb2 = 0.405) |>
  add_row(emb1 = 0.424, emb2 = 0.606) |>
  add_row(emb1 = 0.774, emb2 = 0.405) |>
  add_row(emb1 = 0.774, emb2 = 0.606)

ggplot(data = hex_grid, aes(x = x, y = y)) + 
    geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
    geom_point(data = emb_df, aes(x = emb1, y = emb2), color = "#253494") +
    geom_segment(x = 0.774, y = 0.5055, xend = 0.7, yend = 0.5055,
                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
    geom_segment(x = 0.424, y = 0.5055, xend = 0.35, yend = 0.5055,
                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
    geom_segment(x = 0.599, y = 0.707, xend = 0.599, yend = 0.6,
                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
    geom_segment(x = 0.774, y = 0.405, xend = 0.774, yend = 0.3,
                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
    geom_segment(x = 0.774, y = 0.606, xend = 0.7, yend = 0.55,
                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
    geom_segment(x = 0.6865, y = 0.6565, xend = 0.62, yend = 0.6,
                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
    geom_segment(x = 0.6865, y = 0.3545, xend = 0.6865, yend = 0.25,
                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
  coord_equal()

```

>>>>

```{r}
library(ggplot2)
library(dplyr)
library(quollr)
library(readr)
library(ggforce)

@@

```{r}
#| label: fig-hex-param
#| fig-cap: "Notations for hexagonal grid configurations."

@@

\newpage

@@

Points are assigned to the bin they fall into based on the nearest centroid. If a point is equidistant from multiple centroids, it is assigned to the centroid with the lowest hexagonal bin ID.

```{r}

## To generate embedding data
emb_df <- tibble::tibble(emb1 = 0.37, emb2 = 0.3) |>
  add_row(emb1 = 0.526, emb2 = 1.1) |>
  add_row(emb1 = 0.735, emb2 = 1.1) |>
  add_row(emb1 = 0.631, emb2 = 1.22) |>
  add_row(emb1 = 0.735, emb2 = 1.16) |>
  add_row(emb1 = 0.683, emb2 = 1.19) |>
  add_row(emb1 = 0.735, emb2 = 1.04) |>
  add_row(emb1 = 0.631, emb2 = 0.981) |>
  add_row(emb1 = 0.683, emb2 = 1.0105) |>
  add_row(emb1 = 0.526, emb2 = 1.04) |>
  add_row(emb1 = 0.5785, emb2 = 1.0105) |> 
  add_row(emb1 = 0.526, emb2 = 1.16) |>
  add_row(emb1 = 0.5785, emb2 = 1.19)

## Data to generate circle
circles <- data.frame(
  x = c(0.318, 0.422),
  y = c(0.197,0.378),
  r = rep(0.1205264,2)
)

assign_data_vis <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
  geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 2) +
  geom_point(data = emb_df, aes(x = emb1, y = emb2), color = "#253494", size = 0.5) +
  geom_circle(data = circles, aes(x0 = x, y0 = y, r = r), fill = NA) +
  geom_segment(x = 0.37, y = 0.3, xend = 0.422, yend = 0.378, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.526, y = 1.1, xend = 0.45, yend = 1.1, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.735, y = 1.1, xend = 0.66, yend = 1.1, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.631, y = 1.22, xend = 0.631, yend = 1.15, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.735, y = 1.16, xend = 0.65, yend = 1.12, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.735, y = 1.04, xend = 0.735, yend = 0.95, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.683, y = 1.19, xend = 0.65, yend = 1.12, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.631, y = 0.981, xend = 0.55, yend = 0.95, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.683, y = 1.0105, xend = 0.7, yend = 0.95, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.526, y = 1.04, xend = 0.526, yend = 0.95, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.5785, y = 1.0105, xend = 0.536, yend = 0.97, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.526, y = 1.16, xend = 0.49, yend = 1.12, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.5785, y = 1.19, xend = 0.58, yend = 1.12, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  coord_equal()

```
```{r}
#| label: fig-assign-data
#| fig-cap: "Binning the data. Points are assigned to the nearest centroid. If a point is equidistant from multiple centroids, assigned to the lowest centroid."

assign_data_vis
```
